---
title: "Final Penguins"
author: "Ryan Spake"
date: "12/12/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

Dataset aquired from kaggle. Data is from biologists collecting data on penguins:

https://www.kaggle.com/parulpandey/palmer-archipelago-antarctica-penguin-data?select=penguins_lter.csv
Gorman KB, Williams TD, Fraser WR (2014) Ecological Sexual Dimorphism and Environmental Variability within a Community of Antarctic Penguins (Genus Pygoscelis). PLoS ONE 9(3): e90081. doi:10.1371/journal.pone.0090081


## Prepare the data
Here we are omitting all missing data, Creating dummy variables for island and species, then converting sex into binary.

```{r}
library(caret)
library(class)
library(dplyr)
library(gmodels)
library(knitr)
library(rmarkdown)
library(tidyverse)
library(dplyr)
library(factoextra)
DF_Penguins <- read.csv("C:\\Users\\rspake1\\Desktop\\penguins_lter.csv")
DF_Penguins <- na.omit(DF_Penguins)
Island <- dummyVars(~Island,DF_Penguins)
IslDV <- predict(Island, DF_Penguins)

Species <- dummyVars(~Species,DF_Penguins)
SpecDV <- predict(Species, DF_Penguins)


DF_Penguins <- subset(DF_Penguins, select = -c(Island))
DF_Penguins$Clutch.Completion <- ifelse(DF_Penguins$Clutch.Completion == "Yes",1,0)
DF_Penguins$Sex <- ifelse(DF_Penguins$Sex == "MALE",1,0)

dvDF_Penguins <- cbind(DF_Penguins,IslDV,SpecDV)


clust_constraint <- dvDF_Penguins %>% select_if(is.numeric)
clust_constraint$Sample.Number = NULL



set.seed(2)

clust_constraint <- scale(clust_constraint)
distance <- get_dist(clust_constraint)
fviz_dist(distance)
```


## Finding ideal K

Given that we are seperationg by, islands, k=3 is obvious, though the code below confirms this.

```{r}
clust_constraint <- scale(clust_constraint)
fviz_nbclust(clust_constraint,kmeans, method = "wss")
fviz_nbclust(clust_constraint,kmeans, method = "silhouette")
```


## Visualize clustering -Kmeans

Below we are visually representing the different groups and how they differ statistically.

```{r}
k2 <- kmeans(clust_constraint, centers = 3, nstart = 25)

k2$centers
k2$size
k2$cluster[11]

fviz_cluster(k2, data = clust_constraint)
```


## Identifiy details of clusters
Prepare again

Below we are reestablishing the clusters using clustering.
```{r cars}
library(gmodels)
library(knitr)
library(rmarkdown)
library(readr)
library(tidyverse)
library(caret)
library(cluster)
library(factoextra)
library(RColorBrewer)
library(dplyr)
library(ggraph)
library(igraph)
DF_Penguins <- read.csv("C:\\Users\\rspake1\\Desktop\\penguins_lter.csv")
DF_Penguins <- na.omit(DF_Penguins)

Island <- dummyVars(~Island,DF_Penguins)
IslDV <- predict(Island, DF_Penguins)

Species <- dummyVars(~Species,DF_Penguins)
SpecDV <- predict(Species, DF_Penguins)



DF_Penguins$Clutch.Completion <- ifelse(DF_Penguins$Clutch.Completion == "Yes",1,0)
DF_Penguins$Sex <- ifelse(DF_Penguins$Sex == "MALE",1,0)

dvDF_Penguins <- cbind(DF_Penguins,IslDV,SpecDV)




numeric_Penguins <- dvDF_Penguins %>% select_if(is.numeric)
numeric_Penguins$Sample.Number = NULL


DF_Penguins_norm <- as.data.frame(scale(numeric_Penguins))

d <- dist(DF_Penguins_norm, method = "euclidean")

```


## Select Method

Now we test for which method of clustering will work the best.

```{r pressure, echo=FALSE}
m <- c( "average", "single", "complete", "ward")
names(m) <- c( "average", "single", "complete", "ward")
# function to compute coefficient
ac <- function(x) {
  agnes(d, method = x)$ac
}
map_dbl(m, ac)  
```


## Agnes visualization

Now we develop and Agnes Dendogram to visualize how the clusters separate.
```{r}
hc <- agnes(d, method = "ward")
pltree(hc, cex = 0.6, hang = -1, main = "Agnes Dendogram")
rect.hclust(hc, k = 3, border = 1:5)

```


## Develop cluster partitions

In the two sections below we are creating the cluster partitions, then developing the centroids for each group.
```{r}
cluster_part <- cutree(hc, k = 3)
Penguins_clustered <- mutate(DF_Penguins_norm, cluster = cluster_part)
set.seed(23)

part_index <- createDataPartition(Penguins_clustered$cluster, p = 0.7, list = FALSE)
Part_A <- Penguins_clustered[part_index,]
Part_B <- Penguins_clustered[-part_index,]
```


```{r}
Part_A_centroid <- Part_A %>% gather("features", "values", -cluster) %>% group_by(cluster,features) %>% summarise(mean_values = mean(values)) %>% spread(features, mean_values)

cluster_B <- data.frame(data = seq(1,nrow(Part_B), 1), Cluster_B_Part = rep(0,nrow(Part_B)))

for (x in 1:nrow(Part_B)) {
  cluster_B$Cluster_B_Part[x] <- which.min(as.matrix(get_dist(as.data.frame(rbind(Part_A_centroid[-1], Part_B[x, -length(Part_B)]))))[4,-4])
}

cluster_B <- cluster_B %>% mutate(original_clusters = Part_B$cluster)
mean(cluster_B$Cluster_B_Part) == cluster_B$original_clusters
```


```{r}
split_clusters <- split(Penguins_clustered, Penguins_clustered$cluster)
mean_split <- lapply(split_clusters,colMeans)
mean_split

(centroids <- do.call(rbind, mean_split))
```


## Cluster details

Finally, We are graphing the clusters to identify the details of each cluster. 

```{r}
hc.graph <-
  colorRampPalette(rev(brewer.pal(9, 'Blues')), space = 'Lab')
data.frame(centroids) %>% gather("features", "values",-cluster) %>%
  ggplot(aes(
    x = factor(cluster),
    y = features,
    fill = values
  )) + 
  geom_tile() + theme_classic() +
  theme(
    legend.position = "top",
    plot.title = element_text(hjust = 0.5),
    legend.key.width = unit(3, "cm"),
  ) +
  scale_x_discrete(expand = c(0, 0)) +
  scale_fill_gradientn(colours = hc.graph(100)) +
  labs(title = "Cluster Characteristics",
       x = "Clusters",
       y = "Features",
       fill = "Centroids")
```


## Checking individuals
Below We evaluate k2 so we can Identify which individual Penguin goes in Which cluster. 
```{r}
k2






```

larger the body mass, the larger the beack (culmen) and flipper, but the shallower the culmen.

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
